---
title: "California Coastal Poll"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

## Background

California coastal poll analysis and introduction to R for Coastal and Marine Management class at CSUCI 2019-11.

### Cheat Sheets

- [RStudio IDE](https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf)
- [R Markdown](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf)
- [Data Import](https://github.com/rstudio/cheatsheets/raw/master/data-import.pdf)
- [Data Transformation](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)
- [Data Visualization](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf)

### Notes

**Question 7e Climate change …Thomas Fire?**

- [California wildfires latest: Trump in another war of words over wildfires with California Governor Gavin Newsom - CBS News](https://www.cbsnews.com/news/trump-in-another-war-of-words-over-wildfires-with-california-governor-gavin-newsom/)
- [Opinion | Climate Change in My Backyard (2018-01-11) - The New York Times](https://www.nytimes.com/2018/01/11/opinion/california-floods-mudslides-climate.html?searchResultPosition=1)


**Wide vs long**

  - What uniquely describes an observation?
  - Relationships with joins and relational database tables.

**Data types**

  - numeric
  - character
  - date
  - factor
  - list

## Import Data

```{r read}
# libraries ----
library(glue)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(scales)
library(plotly)
library(DT)

# variables ----
id         <- "1hH68SqNsvAASFn25-X9ssJPkcGSS-SfS6n3zui9hYOQ"
gid        <- "1739121823"
row_beg    <- 4     # first row above data with headers
row_end    <- 1732  # last row of data before summarizing
reload_url <- FALSE

# paths ----
url_csv  <- glue("https://docs.google.com/spreadsheets/d/{id}/export?format=csv&gid={gid}")
all_csv  <- "data/poll_all.csv"
log_csv  <- "data/poll_log.csv"

# download google sheet ----
if (reload_url | !file.exists(all_csv)){
  download.file(url_csv, all_csv)
}

# read headers and data ----
nrows   <- row_end - row_beg

data    <- read_csv(
  all_csv, skip = row_beg - 1, n_max = nrows, guess_max = nrows)

headers <- read_csv(
  all_csv, skip = 0, n_max = row_beg, 
  col_names = F, col_types = strrep("c", ncol(data)))
names(headers) <- names(data)
headers <- headers %>% 
  bind_cols(
    tibble(row = c("question_n", "heading", "question", "answer")))
  
# questions with answers ---
q_a <- headers %>%
  gather(column, value, -row) %>%
  mutate(icol = sapply(1:ncol(data), rep, 4) %>% as.vector()) %>% 
  spread(row, value) %>% 
  arrange(icol) %>%
  fill(question_n) %>% 
  filter(!is.na(answer)) %>%
  mutate(
     question_n = str_remove(question_n, "Question ")) %>% 
  select(question_n, question, heading, answer, icol, column)
#View(q_a)

# questions ----
q <- q_a %>% 
  group_by(question_n) %>% 
  summarize(
    icol     = first(icol),
    question = first(question),
    heading  = first(heading),
    answers  = paste(answer, collapse = " | ")) %>% 
  arrange(icol) %>% 
  select(question_n, question, heading, answers, icol)
# View(q)

# values ----
v <- data %>% 
  filter(str_sub(UID, 1, 8) != "Example ") %>% 
  mutate(UID = as.integer(UID)) %>% 
  gather(column, value, -UID)

# setup list object to pass between functions ----
poll <- list(q=q, q_a=q_a, v=v)

# helper functions ----
question_title <- function(poll, q_n){
  q_i <- filter(poll$q, question_n == q_n)
  glue("{q_i$question_n}. {q_i$question}")
}

plot_question <- function(poll, q_n, answers_rm = c("answered", "comment"), q_rng = c("13")){
  # poll <- list(q=q, q_a=q_a, v=v); q_n <- "7e"; answers_rm = c("answered", "comment")
  # q_n <- "1" # q_n <- "5" # q_n <- "9" # q_n <- "13"
  q_i     <- filter(poll$q, question_n == q_n)
  q_title <- question_title(poll, q_n)
  
  answer_levels <- q_i$answers %>% 
    str_split(fixed(" | "), simplify = T) %>% 
    as.character() %>% 
    setdiff(answers_rm)
  
  if (length(answer_levels) == 0){
    msg <- glue("ERR length(answer_levels) == 0: {q_title}")
    message(msg); return(msg)
    # browser()
  }
  
  v_i <- v %>% 
    left_join(q_a, by="column") %>% 
    filter(question_n == q_n) %>% 
    filter(answer %in% answer_levels) %>% 
    select(UID, answer, value) %>% 
    mutate(
      value  = as.integer(value),
      answer = factor(answer, answer_levels, ordered = T)) %>% 
    filter(!is.na(value))
  
  if (any(na.omit(v_i$value > 1))){
    
    # log questions with values greater than 1
    d_gt1 <- v_i %>% filter(value > 1)
    log_gt1_tbl <- tibble(
      question  = q_title,
      n_values  = nrow(d_gt1),
      condition = "value > 1",
      first5    = glue(
        "UID:   {paste(head(d_gt1$UID,    5), collapse=', ')}; 
        answer: {paste(head(d_gt1$answer, 5), collapse=', ')}; 
        value:  {paste(head(d_gt1$value,  5), collapse=', ')}"))
    if (file.exists(log_csv)){
      write_csv(log_gt1_tbl, log_csv, append=T)
    } else {
      write_csv(log_gt1_tbl, log_csv)
    }
      
    p <- ggplot(v_i, aes(x = answer, y = value, fill = answer)) + 
      geom_boxplot() + 
      theme(legend.position = "none") +
      coord_flip() + 
      ggtitle(q_title) + 
      theme(axis.title = element_blank())
    
  } else {
    
    v_i <- v_i %>% 
      group_by(answer) %>% 
      summarize(
        n   = sum(value, na.rm = T)) %>% 
      mutate(
        pct = n / sum(n))
    
    p <- ggplot(v_i, aes(x = answer, y = pct, fill = answer, label = percent(pct))) + 
      geom_col(position = 'dodge') + 
      geom_text(size = 3) +
      scale_y_continuous(labels = percent) +
      theme(legend.position = "none") +
      coord_flip() + 
      ggtitle(q_title) + 
      theme(axis.title = element_blank())
  }
  print(p)
}

# show questions ----
datatable(q %>% select(-icol), rownames=F)
```

## Poll Answers

```{r visualize, fig.width=6, results="asis"}
# remove questions that are not easy bar charts or box plots
q_n_rm <- c("UID", "Meta", "1")
q_ns <- setdiff(poll$q$question_n, q_n_rm)

# loop over questions
unlink(log_csv) # clear log
for(q_n in q_ns) { # q_n = q_ns[1]
  cat("\n\n###",  question_title(poll, q_n), "\n")
  plot_question(poll, q_n)
}
```

## Log check

Let's look at the log output where values > 1 were detected, which forced the graphing into a boxplot to assume an average value over a range. However where only a few values were > 1, these values might need to be just a 1 or 0 to reflect a Yes or No value to be summed. 

```{r}
read_csv(log_csv) %>% 
  arrange(n_values) %>% 
  datatable()
```



